---
title: "Lab 5 Redo"
author: "Makayla Whitney"
date: "October 21, 2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 
```

1. Create a RStudio Project on your machine.

* Within the Project create a folder for *data* and *scripts*.
* Download the following three datasets into the project *data* folder.
  + *sesame13.sav*
  + *star.csv*
  + *ais.xlsx*

2. Read in the following three datasets, using `{here}` and `{rio}`:
  * *sesame13.sav*
  * *star.csv*
  * *ais.xlsx*
  
**Hint**: For the *ais.xlsx* dataset look at the `skip` argument within the `rio` help documentation.
  
```{r import}
library(rio)
library(here)
library(tidyverse)
library(dplyr)

sesame13 <- import(here("data", "sesame13.sav"))
star <- import(here("data", "star.csv"))
ais <- import(here("data", "ais.xlsx"), skip = 15)
```

3. Using the `ais` data, compute the average red blood cell count and average bmi by sport. Output these data as SPSS and EXCEL files. 

```{r q2}
averages <- ais %>%
  select(rcc, bmi, sport) %>%
  group_by(sport) %>%
  summarize(rcc_avg = mean(rcc), 
            bmi_avg = mean(bmi)) %>%
  as_tibble(ais)

export(averages, here("data", "averages.sav"))
export(averages, here("data", "averages.xlsx"))

import(here("data", "averages.sav"))

```

4. Use the `sesame` dataset to answer the following question: Was the average female age higher in schools or at home? 

```{r q3}

sesame13 %>% 
  characterize() %>% 
  filter(sex == "Female") %>% 
  group_by(setting) %>% 
  summarize(mean_age = mean(age))


```

The average female age was higher in schools than at home. 

5. First, how many rows and columns are in the *star* dataset? Next, remove outliers using a really poor method, just for practice, by eliminating students whose math (*tmathss*) scores were more than three standard deviations **above or below** the corresponding mean. How many rows are in the dataset now?

```{r q4}

star %>%
  filter(tmathss <= mean(tmathss) +3*sd(tmathss) &
           tmathss >= mean(tmathss) -3*sd(tmathss)) %>% 
  nrow()

  
```

6. Use the `star` dataset to compute standardized math and reading scores; name these variables *tmathss* and *treadss*. This means, for each of these variables, subtract the mean from each observation and divide by the standard deviation: $x_s = \frac{x_i - \bar{X}}{sd(X)}$. 
  + Check that your computations were correct by computing the mean and standard deviation of each variable (they should be 0 and 1). 
  + Compute the mean of the standardized variable for all *sex*/*frl* combinations (I'm asking you to extend what you know here, we haven't talked explicitly about how to do this yet, but you have seen examples). 
  + What do you make of these findings? Do you see an effect by *sex*? An *FRL* effect? Is there evidence of an interaction (i.e., that the effect of *FRL* is greater for boys versus girls)?

```{r q5}

star <- star %>%
  mutate(stand_math = (tmathss - mean(tmathss)) / sd(tmathss),
         stand_rdg = (treadss - mean(treadss)) / sd(treadss))

star %>%
  summarize(stand_math_mean = mean(stand_math), 
            stand_math_sd = sd(stand_math),
            stand_rdg_mean = mean(stand_rdg),
            stand_rdg_sd = sd(stand_rdg))

star %>%
  group_by(sex, frl) %>%
  summarize(stand_math_mean = mean(stand_math),
            stand_math_sd = sd(stand_math),
            stand_rdg_mean = mean(stand_rdg),
            stand_rdg_sd = sd(stand_rdg))

```

Please submit you .Rmd file, as well as the rendered .html file.